Local PCA Shows How Population Structure Differs Along the Genome

### 2016-06-06 writeup/local_pca_paper.tex

\begin{figure}
    \begin{center}
       \includegraphics[width=0.9\textwidth]{the-method-diagram}
    \end{center}
    \caption{
         An illustration of the method; see text for details.
         % (1) the genome is divided into windows;
         % (2) population structure is summarized in each with PCA, and 
         % (3) dissimilarities for each are computed;
         % (4) the patterns of dissimilarity are visualized with MDS,
         % and (5) genomic regions clustering together are aggregated and visualized, again with PCA.
         \label{fig:diagram}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
       \includegraphics[width=1\textwidth]{Fig1_Together_MDS_plot}
    \end{center}
    \caption{
         Variation in structure for windows on chromosome arms 2L, 2R, 3L, 3R and X. 
         Each represents a window: the first column shows the MDS visualization of relationships between windows. 
         The second and third columns show the position (midpoint) of each window against the first and second MDS coordinates, respectively. 
         Rows show results from different chromosome arms, from top to bottom: Chr2L, Chr2R, Chr3L, Chr3R, ChrX. 
         Colors are consistent for plots in each row. 
         Vertical black lines show the breakpoints of known polymorphic inversions for Chr2L, Chr2R, Chr3L, Chr3R (from top to down: In(2L)t, In(2R)NS, In(3L)OK, In(3R)K).
         \label{fig:mds_chr2L}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
       \includegraphics[width=0.9\textwidth]{Fig2_all_pca_plots}
    \end{center}
    \caption{      
        PCA plots for the three sets of genomic windows colored in Figure \ref{fig:mds_chr2L}. 
        Each point represents a sample. 
        The plots are plotting the second principal components against the first principal components. 
        The first column shows the windows colored green in Figure 1; the second column shows orange windows; and third shows purple windows.
         \label{fig:pca_chr2L}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
       \includegraphics[width=0.9\textwidth]{Fig3}
    \end{center}
    \caption{
         As in Figure 2, except that samples are colored by orientation of the corresponding polymorphic inversion, In(2L)t, In(2R)NS, In(3L)OK and In(3R)K. (Data from \citet{lack2015drosophila}.
        \label{fig:color_inver}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
       \includegraphics[width=0.9\textwidth]{Fig4_temp}
    \end{center}
    \caption{
         Variation in structure between windows on human chromosomes 8, 15, and 17. 
         Each point in the plot represents a window.
         The first column shows the MDS visualization of relationships between windows;
         the second and third columns show the two MDS coordinates of each window against its position (midpoint) along the chromosome. 
         Rows, from top to bottom show chromosomes 8, 15, and 17. 
         The vertical red lines show the breakpoints of the known inversions \citep{antonacci2009characterization}.
        \label{fig:mds_human}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
       \includegraphics[width=0.7\textwidth]{Medicago_MDS_1D_and_gene_count}
    \end{center}
    \caption{
         MDS results and gene density for the \textit{Medicago} genome (chromosome 1-8). 
         For each chromosome, the red plot above is first coordinate of MDS against the middle position of each window along each chromosome. 
         The black plot below is gene count for each window against the position of each window.
         \label{fig:mds_medicago}
    }
\end{figure}

\begin{figure}
    \begin{center}
       \includegraphics[width=0.9\textwidth]{medicago_mds}
    \end{center}
    \caption{
     MDS visualization for Medicago chromosome 3. Each point in the plot stands for a window (length 1000 SNPs).
       \label{fig:mds12_medicago}
    }
\end{figure}

\begin{figure}
    \begin{center}
       \includegraphics[width=0.9\textwidth]{medicago_peaks_pca}
    \end{center}
    \caption{
        PCA plots for the three sets of genomic windows colored separately in Figure \ref{fig:mds12_medicago}. 
        From left to right, the windows colored green, orange, and purple. Each point corresponds to a sample, colored by origin.
        \label{fig:pca_medicago}
    }
\end{figure}

\begin{figure}
    \begin{center}
       \includegraphics[width=0.9\textwidth]{Fig6_MDS_1D_against_gene_count_all_chr_win104_with_lm}
    \end{center}
    \caption{
        First MDS coordinate against gene density for all 8 chromosomes. 
        The first MDS coordinate is significantly correlated with gene count (r=0.146, p=$2.2\times 10^{-16}$). 
        (See supplementary for each single chromosome's MDS result against gene count for \textit{Medicago})
        \label{fig:mds_gene_count}
    }
\end{figure}

>>>>

\begin{figure}
    \begin{center}
       \includegraphics{the-method-diagram}
    \end{center}
    \caption{
         An illustration of the method; see text for details.
         % (1) the genome is divided into windows;
         % (2) population structure is summarized in each with PCA, and 
         % (3) dissimilarities for each are computed;
         % (4) the patterns of dissimilarity are visualized with MDS,
         % and (5) genomic regions clustering together are aggregated and visualized, again with PCA.
         \label{fig:diagram}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
       \includegraphics{Fig1_allchr_Together_MDS_plot_compact}
    \end{center}
    \caption{
         Variation in structure for windows on chromosome arms 2L, 2R, 3L, 3R and X. 
         Each represents a window: the first column shows the MDS visualization of relationships between windows. 
         The second and third columns show the position (midpoint) of each window against the first and second MDS coordinates, respectively. 
         Rows show results from different chromosome arms, from top to bottom: Chr2L, Chr2R, Chr3L, Chr3R, ChrX. 
         Colors are consistent for plots in each row. 
         Vertical black lines show the breakpoints of known polymorphic inversions for Chr2L, Chr2R, Chr3L, Chr3R (from top to down: In(2L)t, In(2R)NS, In(3L)OK, In(3R)K).
         \label{fig:mds_chr2L}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
       \includegraphics{Fig2_pca_plots_allchr_3peaks_byMDS}
    \end{center}
    \caption{      
        PCA plots for the three sets of genomic windows colored in Figure \ref{fig:mds_chr2L}. 
        Each point represents a sample. 
        The plots are plotting the second principal components against the first principal components. 
        The first column shows the windows colored green in Figure 1; the second column shows orange windows; and third shows purple windows.
         \label{fig:pca_chr2L}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
       \includegraphics{Fig3_pca_plots_color_by_inv_allchr}
    \end{center}
    \caption{
         As in Figure 2, except that samples are colored by orientation of the corresponding polymorphic inversion, In(2L)t, In(2R)NS, In(3L)OK and In(3R)K. (Data from \citet{lack2015drosophila}.
        \label{fig:color_inver}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
       \includegraphics{Fig4_temp}
    \end{center}
    \caption{
         Variation in structure between windows on human chromosomes 8, 15, and 17. 
         Each point in the plot represents a window.
         The first column shows the MDS visualization of relationships between windows;
         the second and third columns show the two MDS coordinates of each window against its position (midpoint) along the chromosome. 
         Rows, from top to bottom show chromosomes 8, 15, and 17. 
         The vertical red lines show the breakpoints of the known inversions \citep{antonacci2009characterization}.
        \label{fig:mds_human}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
       \includegraphics{Medicago_MDS_1D_and_gene_count}
    \end{center}
    \caption{
         MDS results and gene density for the \textit{Medicago} genome (chromosome 1-8). 
         For each chromosome, the red plot above is first coordinate of MDS against the middle position of each window along each chromosome. 
         The black plot below is gene count for each window against the position of each window.
         \label{fig:mds_medicago}
    }
\end{figure}

\begin{figure}
    \begin{center}
       \includegraphics{medicago_mds}
    \end{center}
    \caption{
     MDS visualization for Medicago chromosome 3. Each point in the plot stands for a window (length 1000 SNPs).
       \label{fig:mds12_medicago}
    }
\end{figure}

\begin{figure}
    \begin{center}
       \includegraphics{medicago_peaks_pca}
    \end{center}
    \caption{
        PCA plots for the three sets of genomic windows colored separately in Figure \ref{fig:mds12_medicago}. 
        From left to right, the windows colored green, orange, and purple. Each point corresponds to a sample, colored by origin.
        \label{fig:pca_medicago}
    }
\end{figure}

\begin{figure}
    \begin{center}
       \includegraphics{Fig6_MDS_1D_against_gene_count_all_chr_win104_with_lm}
    \end{center}
    \caption{
        First MDS coordinate against gene density for all 8 chromosomes. 
        The first MDS coordinate is significantly correlated with gene count (r=0.146, p=$2.2\times 10^{-16}$). 
        (See supplementary for each single chromosome's MDS result against gene count for \textit{Medicago})
        \label{fig:mds_gene_count}
    }
\end{figure}
